Equations & R Code

### 2017-04-18 manuscript/coexistence-stability_SI.Rmd

  if(Rsd_annual == 0) {
    Rvector        <- rlnorm(seasons, Rmu, Rsd_annual)
  }
  if(Rsd_annual > 0) {
    Rvector        <- urlnorm(seasons, Rmu, Rsd_annual, lb = 0, ub = 200)
  }

@@

\section{Eight-Species Storage Effect Model}
In the main text we constrained our focus to four-species communities because getting more than four species to coexist by relative nonlinearity is tricky, and usually requires adding another coexistence mechanism on top off relative nonlinearity [@Yuan2015].
The storage effect does not suffer from this limitation, but we wanted our results in the main text to be easily comparable between coexistence mechanisms.
Here, we show that our results are qualitatively similar if we simulate an eight-species community with species coexistence maintained by the storage effect.
We conducted the same numerical simulations described in the main text for Figure 2.
Quoting from the main text:

@@

1. Environment-competition covariance (*EC* covariance) becomes less positive
2. Environemntal variation decreases
3. Buffering of population growth rate declines

@@

Our results are qualitatively similar: (1) increasing environmental variation allows more species to coexist, creating a positive relationship between species richness and ecosystem variability (richness increases along x-axes within panels); (2) and at any given level of environmental variability, it is always better to have more species, creating a within-site negative relationship between species richness and ecosystem variability (Fig. SI-2).
Although the qualitative patterns are similar, increasing the dormant mortality rate has two interesting effects: (1) it makes coexistence more difficult, (2) it reduces the absolute value of ecosystem *CV* by reducing mean population size a little bit and by reducing the temporal standard deviation a lot, and (3) it weakens the buffering effect of additional species (compare the spread of the lines between top and bottom panels of Fig. SI-2).
The second result occurs because population fluctuations are reduced when very little biomass can be activated to the live stage at the beginning of each season.
That is, in terms of an annual plant, germination of live biomass is always a fraction of a very small number.
The third result is a consequence of the second: populations are not fluctuating that much, so total *CV* can only be reduced by so much with the addition of another species.

@@

Condition 1 is already included in our main analysis (Figures 2 and 4 in the main text).
Condition 2 simply creates conditions where some species cannot coexist, which would not change the relationship between species richness and ecosystem variability, but rather disallow it.
Condition 3 is more interesting, because it could create situations where species may not coexist, but it could also weaken the diversity--ecosystem variabilty relationship be reducing mean biomass and the effect of species with very nonlinear growth rates.
To explore the effect of condition 3, we ran simulations across a gradient of resource variability (11 levels) crossed with a gradient of mean resource levels (4 levels: $\mu_R = 1,2,3,4$).

As expected, lower mean levels of resource make coexistence more difficult because resource acquisitive species are never favored (Fig. SI-3).
Our general conclusion remains, however, that positive diversity--ecosystem variability relationships are possible.
However, the occurence a of positive diversity--ecosystem variability relationship depends on amount of resource variability being sampled over and the traits of individual species (e.g., resource uptake curves). 
In other words, the patterns we report in the main text require that fluctuation-dependent species coexistence be possible given the environmental conditions and species' traits.

\begin{figure}[!ht]
  \centering
      \includegraphics[width=6in]{./components/SI_relative_nonlinearity_four_rmus.png}
  \caption{The effect of environmental variability on ecosystem variability when species coexist via relative nonlinearity. The species pool increases from one to four species, with the fourth species being most unstable (e.g., resource conservative to resource acquisitive). Increasing environmental variability (the SD of annual resource availability) allows for greater species richness, but species additions do not modulate the effect of environmental variability on ecosystem variability. Higher values of mean resource availability (left to right across panels) makes coexistence easier. Parameter values are as in Figure 2C in the main text.}
\end{figure}

>>>>

  if(Rsd_annual == 0) { # urlnorm can't handle zero variance
    Rvector        <- rlnorm(seasons, Rmu, Rsd_annual)
  }
  if(Rsd_annual > 0) { # when variance greater than zero
    Rvector        <- urlnorm(seasons, Rmu, Rsd_annual, lb = 0, ub = 200)
  }

@@

\section{Eight-Species Storage Effect Model}
In the main text we constrained our focus to four-species communities because getting more than four species to coexist by relative nonlinearity is tricky, and usually requires adding another coexistence mechanism on top of relative nonlinearity [@Yuan2015].
The storage effect does not suffer from this limitation, but we wanted our results in the main text to be easily comparable between coexistence mechanisms.
Here, we show that our results are qualitatively similar if we simulate an eight-species community with species coexistence maintained by the storage effect.
We conducted the same numerical simulations described in the main text for Figure 2.
Quoting from the main text:

@@

1. Environment-competition covariance (*EC* covariance) becomes less positive
2. Environmental variation decreases
3. Buffering of population growth rate declines

@@

Our results are qualitatively similar: (1) increasing environmental variation allows more species to coexist, creating a positive relationship between species richness and ecosystem variability (richness increases along x-axes within panels); (2) and at any given level of environmental variability, it is always better to have more species, creating a within-site negative relationship between species richness and ecosystem variability (Fig. SI-2).
Although the qualitative patterns are similar, increasing the dormant mortality rate has three consistent effects: (1) it makes coexistence more difficult, (2) it reduces the absolute value of ecosystem *CV* by slightly reducing mean population size and significantly reducing temporal standard deviation, and (3) it weakens the buffering effect of additional species (compare the spread of the lines between top and bottom panels of Fig. SI-2).
The second result occurs because population fluctuations are reduced when very little biomass can be activated to the live stage at the beginning of each season.
That is, in terms of an annual plant, germination of live biomass is always a fraction of a very small number.
The third result is a consequence of the second: populations are not fluctuating that much, so total *CV* can only be reduced by as much with the addition of another species.

@@

Condition 1 is already included in our main analysis (Figures 2 and 4 in the main text).
Condition 2 simply creates conditions where some species cannot coexist, which would not change the relationship between species richness and ecosystem variability, but rather disallow it.
Condition 3 could create situations where species may not coexist, but it could also weaken the diversity--ecosystem variabilty relationship by reducing mean biomass and the effect of species with very nonlinear growth rates.
To explore the effect of condition 3, we ran simulations across a gradient of resource variability (11 levels) crossed with a gradient of mean resource levels (4 levels: $\mu_R = 1,2,3,4$).

As expected, lower mean levels of resource make coexistence more difficult because resource acquisitive species are never favored (Fig. SI-3).
Our general conclusion remains, however, that positive diversity--ecosystem variability relationships are possible.
However, the occurrence of a positive diversity--ecosystem variability relationship depends on the extent of the resource variability gradient and the traits of individual species (e.g., resource uptake curves). 
In other words, the patterns we report in the main text require that fluctuation-dependent species coexistence be possible given the environmental conditions and species' traits.

\begin{figure}[!ht]
  \centering
      \includegraphics[width=7in]{./components/SI_relative_nonlinearity_four_rmus.png}
  \caption{The effect of environmental variability on ecosystem variability when species coexist via relative nonlinearity. The species pool increases from one to four species, with the fourth species being most unstable (e.g., resource conservative to resource acquisitive). Increasing environmental variability (the SD of annual resource availability) allows for greater species richness, but species additions do not modulate the effect of environmental variability on ecosystem variability. Higher values of mean resource availability (left to right across panels) makes coexistence easier. Parameter values are as in Figure 2C in the main text.}
\end{figure}

### 2017-04-18 manuscript/coexistence-stability_ms.tex

\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\usepackage{todonotes}
\doublespacing
\usepackage{rotating}
\usepackage{color, soul}
\usepackage[font={normalsize},labelfont={bf},labelsep=quad]{caption}
\usepackage{sectsty}
\usepackage{bm,mathrsfs}
\usepackage{mathptmx}

@@

@@

When we allowed communities to assemble at sites across a gradient of
environmental variability, we discovered a positive relationship between
species richness and ecosystem variability (Fig. 2A,C). While surprising
when viewed through the lens of biodiversity--ecosystem functioning
theory and experimental findings, such a relationship is predicted by
theory on coexistence in fluctuating environments. Environmental
variability is a prerequisite for the storage effect and relative
nonlinearity to stabilize coexistence (Chesson 2000a). These mechanisms
can translate increased variability into higher species richness (Fig.
SI-6), but the increase in environmental variability also increases
ecosystem variability.
\new{However, the apparent saturation of the relationship in Fig. 3A suggests that the portfolio effects that buffer ecosystems against environmental variability, and inherently emerge under the storage effect, get stronger as more species are able to coexist.
Indeed, the relationship between species richness and ecosystem \emph{CV} completely saturates under the storage effect in more species rich communities (Fig. SI-1).
This suggests neutral diversity--ecosystem variability relationships are possible due to the storage effect.}

>>>>

\usepackage{lineno}
\usepackage{setspace}
\usepackage{todonotes}
\doublespacing
\usepackage{rotating}
\usepackage{color, soul}
\usepackage[font={normalsize},labelfont={bf},labelsep=quad]{caption}
\usepackage{sectsty}
\usepackage{bm,mathrsfs}
\usepackage{mathptmx}

@@

\linenumbers

@@

When we allowed communities to assemble at sites across a gradient of
environmental variability, we discovered a positive relationship between
species richness and ecosystem variability (Fig. 2A,C). While surprising
when viewed through the lens of biodiversity--ecosystem functioning
theory and experimental findings, such a relationship is predicted by
theory on coexistence in fluctuating environments. Environmental
variability is a prerequisite for the storage effect and relative
nonlinearity to stabilize coexistence (Chesson 2000a). These mechanisms
can translate increased variability into higher species richness (Fig.
SI-6), but the increase in environmental variability also increases
ecosystem variability.
\new{However, the apparent saturation of the relationship in Fig. 2A suggests that the portfolio effects that buffer ecosystems against environmental variability, and inherently emerge under the storage effect, get stronger as more species are able to coexist.
Indeed, the relationship between species richness and ecosystem \emph{CV} completely saturates under the storage effect in more species rich communities (Fig. SI-1).
This suggests neutral diversity--ecosystem variability relationships are possible due to the storage effect.}
